---
title: "Assignment #5, Initial Data Analysis"
author: "YOUR NAME"
date: 'Updated on `r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
     theme: journal
     highlight: textmate
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 999)
```

### Load Libraries

```{r libraries, warning = FALSE, message = FALSE}
library(tidyverse)
library(sf)
library(tmap)
library(knitr)
library(kableExtra)
```

### Read Data

```{r readdata, warning = FALSE, message = FALSE}
# Read NC Zip Code COVID data
nc.zip.covid <- st_read("../Data/NC_ZIP_COVID.shp", 
                        quiet = TRUE)

```

### Descriptive Statistics

```{r summarize, warning = FALSE, message = FALSE}
# Create summary information holder
# for Deaths per 1000 residents information
summary.table <- tibble(Measure = c("Observations",
                                    "Minimum",
                                    "Maximum",
                                    "Mean",
                                    "Median",
                                    "Standard Deviation"),
                        `Cumulative Cases` = c(sum(!is.na(nc.zip.covid$CASES)),
                                       min(nc.zip.covid$CASES, na.rm = TRUE),
                                       max(nc.zip.covid$CASES, na.rm = TRUE),
                                       mean(nc.zip.covid$CASES, na.rm = TRUE),
                                       median(nc.zip.covid$CASES, na.rm = TRUE),
                                       sd(nc.zip.covid$CASES, na.rm = TRUE)
                                       ),
                        `Case Rate` = c(sum(!is.na(nc.zip.covid$CASERATE)),
                                      min(nc.zip.covid$CASERATE, na.rm = TRUE),
                                      max(nc.zip.covid$CASERATE, na.rm = TRUE),
                                      mean(nc.zip.covid$CASERATE, na.rm = TRUE),
                                      median(nc.zip.covid$CASERATE, na.rm = TRUE),
                                      sd(nc.zip.covid$CASERATE, na.rm = TRUE)
                                      )
                        )

## Print nice version of the table
kable(summary.table, 
      digits = 1,
      format.args = list(big.mark = ",",
                         scientific = FALSE,
                         drop0trailing = TRUE),
      caption = "Summary of North Carolina COVID-19 Cumulative Cases and Case Rates (per 1,000 residents) by Zip Code") %>% 
  kable_styling(bootstrap_options = c("striped", 
                                      "hover", 
                                      "condensed", 
                                      "responsive"), 
                full_width = F)


```


### Histogram 1

```{r histogram1, warning = FALSE, message = FALSE, fig.width = 7, fig.height = 4}
## plot histogram
ggplot(nc.zip.covid, 
       aes(x = CASERATE)) + 
  geom_histogram(binwidth = 0.5) +
  xlab("Cumulative COVID-19 cases per 1,000 People") +
  ggtitle("Histogram of COVID-19 Case Rates") +
  theme_minimal()

```


### Histogram 2

```{r histogram2, warning = FALSE, message = FALSE, fig.width = 7, fig.height = 4}
## plot histogram
ggplot(nc.zip.covid, 
       aes(x = CASERATE)) + 
  geom_histogram(binwidth = 2) +
  xlab("Cumulative COVID-19 cases per 1,000 People") +
  ggtitle("Histogram of COVID-19 Case Rates") +
  theme_minimal()

```

### Histogram 3

```{r histogram3, warning = FALSE, message = FALSE, fig.width = 7, fig.height = 4}
## plot histogram
ggplot(nc.zip.covid, 
       aes(x = CASES)) + 
  geom_histogram(binwidth = 50) +
  xlab("Cumulative COVID-19 cases") +
  ggtitle("Histogram of COVID-19 Cases") +
  theme_minimal()

```

### Choropleth Map 1

```{r map1, warning = FALSE, message = FALSE, fig.width = 9, fig.height = 5}
## Plot map
tm_shape(nc.zip.covid) + 
  tm_polygons("CASES", 
              style = "jenks", 
              palette = "YlOrRd",
              lwd = 0.25,
              border.col = "black",
              border.alpha = 0.5,
              title = "Cumulative COVID-19 Cases")
```


### Choropleth Map 2

```{r map2, warning = FALSE, message = FALSE, fig.width = 9, fig.height = 5}
## Plot map
tm_shape(nc.zip.covid) + 
  tm_polygons("CASERATE", 
              style = "jenks", 
              palette = "YlOrRd",
              lwd = 0.25,
              border.col = "black",
              border.alpha = 0.5,
              title = "Cumulative COVID-19 Case Rate (per 1,000 people)")
```


### Choropleth Map 3

```{r map3, warning = FALSE, message = FALSE, fig.width = 9, fig.height = 5}
## Plot map
tm_shape(nc.zip.covid) + 
  tm_polygons("CASERATE", 
              style = "quantile", 
              palette = "YlOrRd",
              lwd = 0.25,
              border.col = "black",
              border.alpha = 0.5,
              title = "Cumulative COVID-19 Case Rate (per 1,000 people)")
```

******

##### Question 1. Write one sentence that **uses information from the table above** to describe the central tendency and dispersion of Cumulative COVID-19 Case Rates (per 1,000 people) by Zip Code. (1 sentence, using at least one measure of central tendency and one measure of dispersion)

YOUR ANSWER HERE  

  
******


##### Question 2. What is the difference between Histogram 1 and Histogram 2?  Which one of these is more effective and why? (2-3 sentences)

YOUR ANSWER HERE  

  
******


##### Question 3. What are the differences between the data distributions found in Histogram 2 and Histogram 3? (2-3 sentences)

YOUR ANSWER HERE  

  
******


##### Question 4. Why is Choropleth Map 2 more appropriate for mapping and analysis than Choropleth Map 1? (2-3 sentences) 

YOUR ANSWER HERE  

  
******


##### Question 5. Describe the nature of the *spatial pattern* found in Choropleth Map 2. (1-2 sentences) 

YOUR ANSWER HERE  

  
******


##### Question 6. Explain the difference between Choropleth Map 3 and Choropleth Map 2. Which is easier for you to interpret? (2-3 sentences) 

YOUR ANSWER HERE  

  
******