---
title: "Holmes_GEOG215_Final"
author: "Emma Holmes"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  html_document:
    theme: "cosmo"
    highlight: "textmate"
---

```{r setup, include = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(tmap)
library(knitr)
library(kableExtra)
library(sp)
library(spdep)
```

## Introduction
I analyzed trees in London, UK by comparing the location of the trees that are in the different boroughs of Greater London. This is useful information for a variety of people to know if there is a lack of trees in any given area. Then with this information there can be an effort to plant more trees in places with less of them. Trees are important in an area's quality of life as they filter the nearby air and water, including removing carbon, and act as cooling agents. 

## Data Preparation

The dataset of tree locations was found on the London Datastore (https://data.london.gov.uk/dataset/local-authority-maintained-trees) and was collected by the Greater London Authory. 

The shapefile of the boroughs in Greater London was found on the European Commission's website (https://data.europa.eu/data/datasets/statistical-gis-boundary-files-for-london/?locale=en).

To prepare my data for the analysis, I first added geometry to the tree locations using the longitude and latitude that had been provided. Then I transformed the CRS to be 27700, which is the local CRS for London that the borough data is already in.

```{r, include = FALSE}
tree.loc <- read.csv("Data/Borough_tree_list_2021.csv")

boroughs <- st_read("Data/statistical-gis-boundaries-london/ESRI/London_Borough_Excluding_MHW.shp")

```

```{r, warning = FALSE}

## Make trees numeric
num.trees <- tree.loc %>%
  mutate_at(12:13, as.numeric)

## One point got messed up and will skew the data (51.585, 20180214)
## Remove this point from the data
cl.trees <- num.trees %>%
  filter(!row_number() %in% c("764094"))

## Get rid of NAs and blanks in the data
cn.trees <- cl.trees[,c("lon","lat","borough")] %>% 
  na_if("") %>% # recode empty strings "" by NAs
  na.omit() # remove NAs

## Make trees spatial
trees_sf <- cn.trees %>%
  st_as_sf(coords = c("lon","lat"), remove = FALSE)

## Add CRS
st_crs(trees_sf) = 4326

## Transform Tree Location CRS to OSGB36 (ESPG = 27700)
trees_sf = st_transform(trees_sf, 27700)

## Make shapefile valid
boroughs <- st_make_valid(boroughs)

## Remove extra columns in boroughs
cl.boroughs <- boroughs %>%
  select(!c("HECTARES","NONLD_AREA","ONS_INNER","SUB_2009","SUB_2006"))

```

## Analysis

### Data Description and Summary

Looking at the table below, it is already apparent that there is a large range between the number of trees that are in each borough as the maximum is 77,589 and the minimum is 252. The median is also smaller than the mean which tells me that the data is skewed to the right, with the majority of the boroughs having a smaller amount of trees.

```{r, include = FALSE}
## Create a data frame for the table

#Sum of the trees per borough
tree.per.borough <- trees_sf %>% 
  count(borough)

#Total amount of trees in London
cal.trees <- sum(tree.per.borough$n)

#Mean amount of trees per borough
cal.trees[2] <- mean(tree.per.borough$n)

#Median amount of trees per borough
cal.trees[3] <- median(tree.per.borough$n)

#Minimum amount of trees per borough
cal.trees[4] <- min(tree.per.borough$n)

#Maximum amount of trees per borough
cal.trees[5] <- max(tree.per.borough$n)

#Add names
names(cal.trees)[1:5] <- c("Sum", "Mean", "Median", "Minimum", "Maximum")

# Convert to dataframe
tree.bor.dat <- data.frame("Metric" = names(cal.trees), 
                       "Value" = cal.trees)

## Create the table
```

```{r printtable, echo = FALSE, warning = FALSE}
kable(tree.bor.dat, 
      digits = 3,
      format.args = list(big.mark = ",",
                         scientific = FALSE,
                         drop0trailing = TRUE),
      caption = "Statistics of Trees Per Borough") %>% 
  kable_styling(bootstrap_options = c("striped", 
                                      "hover", 
                                      "condensed", 
                                      "responsive"), 
                full_width = F)
```

### Geographic Distribution and Spatial Clustering

##### Barchart
This barchart gives a visualization of the amount of trees in each of the boroughs with two of the boroughs having over 75,000 trees, but the majority of the boroughs fall below having 40,000 trees. It reiterates the how the data is skewed towards lower amount of trees in each borough.

```{r barplot, echo = FALSE, warning = FALSE, fig.width = 9, fig.height = 6}
## Create a barchart of the number of trees in each borough

# Get table information from sp data
t.tree.per.borough <- st_drop_geometry(tree.per.borough)

# Sort data by Cases (high to low)
t.tree.per.borough <- t.tree.per.borough[order(t.tree.per.borough$n, decreasing = TRUE), ]

## Create ggplot
ggplot(t.tree.per.borough,
       aes(x = borough,
           y = n)) + 
  geom_bar(stat = "identity") + 
  xlab("Borough") +
  ylab("Number of Trees") +
  coord_flip() +
  theme_minimal() +
  ggtitle("Trees per Borough") +
  scale_y_continuous(breaks=c(5000, 10000, 15000, 20000, 25000, 30000, 35000, 40000, 45000, 50000, 55000, 60000, 65000, 70000, 75000, 80000))

```

##### Choropleth Map

The choropleth shows that there are more trees reported in the northern part of Greater London and less trees reported in the southern part of Greater London.

```{r map1, echo = FALSE, warning = FALSE, fig.width = 9, fig.height = 6}
## Create a map plotting the trees onto the heat map with the boroughs

# Spatial join the trees per boroughs and the boroughs
# Confirm CRS
st_crs(cl.boroughs) <- 27700
st_crs(tree.per.borough) <- 27700

tr.bo_sf <- st_join(cl.boroughs,
                   tree.per.borough,
                   join = st_intersects)

# Make the choropleth interactive
tmap_mode("view")

#Plot the choropleth map
tm_shape(tr.bo_sf)+
  tm_polygons("n",
              style = "jenks", 
              palette = "Greens",
              lwd = 0.25,
              border.col = "black",
              border.alpha = 0.5,
              title = "Number of Trees Per Borough")

```

##### Moran's I

Moran's I is 0.049. The p-value is 0.00175.

```{r}
## Use Moran's I to analyze whether the data is more clustered or dispersed.

# Create Queen case neighbors
tr.bo_queen <- poly2nb(tr.bo_sf, 
                       queen = TRUE)

# Convert neighbors to weight matrix
tr.bo_weights <- nb2listw(tr.bo_queen, 
                          style = "B",         # B is binary (1,0)
                          zero.policy = TRUE)  # zero.policy allows for observations with NO neighbors 

# Moran's I
tr.bo.moran <- moran.test(tr.bo_sf$n, # The column in your sp data
                             tr.bo_weights, # Weights object
                             zero.policy = TRUE, # Allows for observations with NO neighbors
                             randomisation = TRUE) # Compares to randomized NULL data

# Print Moran's I
tr.bo.moran

```

##### Map of Local Indicator of Spatial Association

The LISA Map shows one cluster with high values and the neighbors also having high values in the north of Greater London, and a smaller area with low values where neighbors also had low values.

The summary report of the LISA states that there are a couple areas of low values surrounded by high valued neighbors and high values surrounded by low value neighbors, but the map does not show these areas so they are not strong enough to be relevant.

```{r map2, echo = FALSE, warning = FALSE, fig.width = 9, fig.height = 6}
## Use LISA to analyze the clustering of trees throughout London

# LISA -- Local Moran's I
tr.bo.lisa <- localmoran(tr.bo_sf$n, # The column in your sp data 
                         tr.bo_weights, # Weights object
                         zero.policy = TRUE) %>% # Best to keep TRUE for LISA
  as.data.frame() # Make result into data frame

# Scale the input data to deviation from mean
cDV <- tr.bo_sf$n - mean(tr.bo_sf$n) 

# Get spatial lag (neighbors) values for each observation
lagDV <- lag.listw(tr.bo_weights, tr.bo_sf$n)

# Scale the lag values to deviation from mean
clagDV <- lagDV - mean(lagDV, na.rm = TRUE)

# Add holder column with all 0s
tr.bo.lisa$Cat <- rep("0", nrow(tr.bo.lisa))

# Add a label based on the values
tr.bo.lisa$Cat[which(cDV > 0 & clagDV > 0 & tr.bo.lisa[,5] < 0.05)] <- "HH" 
tr.bo.lisa$Cat[which(cDV < 0 & clagDV < 0 & tr.bo.lisa[,5] < 0.05)] <- "LL"      
tr.bo.lisa$Cat[which(cDV < 0 & clagDV > 0 & tr.bo.lisa[,5] < 0.05)] <- "LH"
tr.bo.lisa$Cat[which(cDV > 0 & clagDV < 0 & tr.bo.lisa[,5] < 0.05)] <- "HL"

# Quick summary of LISA output
table(tr.bo.lisa$Cat)

# Add LISA category column to the spatial data
tr.bo_sf$LISACAT <- tr.bo.lisa$Cat

# Plot the map
tm_shape(tr.bo_sf) + 
  tm_polygons("LISACAT", 
              style = "cat", 
              palette = c("grey", 
                          "red", 
                          "pink", 
                          "lightblue", 
                          "blue"), 
              border.col = "Black", 
              border.alpha = 0.25) +
  tm_layout(legend.outside = TRUE)

```

### Interpretation

Moran's I, 0.049, is closer to 1 than -1 so it shows very slight indication of clustering but it is barely above 0 so it indicates the data is random. And the p-value, 0.00175, is smaller than the alpha level of 0.05 which indicates that Moran's I is statistically significant. The data is more random and there is no consistent pattern of clustering or dispersing.

The choropleth LISA map show a cluster of a large amount of trees in the north of Greater London, while there is a lack of trees near the City of London. The cluster of low values at the City of London makes sense because it has less green space. Currently I do not know why there is a cluster of high values near Enfield or Waltham Forest (other than the name including Forest), but further research could be done to find out the reason behind this, perhaps these areas are more residential or have wealthier inhabitants.

The primary limitation to this data that there is a lack of consistency among the data of tree locations provided. Many of the tree locations are from 2021, but some boroughs had not submitted updated numbers so they are from 2018 or 2015/2016.

## Conclusion

The data analyzed shows that there is not a pattern of clustering or dispersing of the total amount of tree locations in each borough. To me this shows that a few boroughs have more trees while others are lacking trees. Further research can be done on whether or not the lack of trees in these other boroughs has any affect on their quality of living, an example being whether the lack of trees increases the amount of heat felt in the summer.
