---
title: "Assignment #4"
author: "Emma Holmes"
date: 'Updated on `r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
     theme: spacelab
     highlight: textmate
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 999)
library(tidyverse)
library(sf)
library(tmap)
library(knitr)
library(kableExtra)
library(here)
```

```{r readdata, include = FALSE}
# Read NC state-level data
nc.st.dat <- read_csv(here("Output/NC_State_COVID_POP_2020-09-23.csv"))

# Read NC zip-level data (for mapping)
nc.zip.dat <- read_csv(here("Output/NC_ZIPmapping_COVID_POP_2020-09-23.csv"))

# Read NC zip spatial data layer (simplified for mapping)
nc.zip.sp <- st_read(here("Output/NC_ZIP_2020_mapping.shp"))

# Join/merge spatial and table data
nc.zip.sp <- merge(nc.zip.sp, 
                   nc.zip.dat, 
                   by.x = "ZIP_CODE",
                   by.y = "ZIPCODEmap",
                   all.x = TRUE)
```

******

### State-level table

In this part of the document, we will use `kable` to print a nice table of the State-level results created in the file GEOG215-data-analysis.R. We won't show the code, just the table!

```{r printtable, echo = FALSE, warning = FALSE}
kable(nc.st.dat, 
      digits = 3,
      format.args = list(big.mark = ",",
                         scientific = FALSE,
                         drop0trailing = TRUE),
      caption = "North Carolina COVID-19 Statistics") %>% 
  kable_styling(bootstrap_options = c("striped", 
                                      "hover", 
                                      "condensed", 
                                      "responsive"), 
                full_width = F)
```

We can also use inline code to write statements such as: There have been `r format(nc.st.dat$Value[2], big.mark = ",")` lab confirmed COVID-19 cases in North Carolina, and, sadly, `r format(nc.st.dat$Value[3], big.mark = ",")` people have died.

******

### Ordered Bar Chart (top 20 Zip Codes, Cases)

Now, we will use `ggplot` to create a nice graphic showing the name (and case count) of the 20 Zip Codes with the highest case counts.

```{r barplot, echo = FALSE, warning = FALSE, fig.width = 7, fig.height = 4}
## Get table information from sp data
nc.zip.sp.t <- st_drop_geometry(nc.zip.sp)

## Sort data by Cases (high to low)
nc.zip.sp.t <- nc.zip.sp.t[order(nc.zip.sp.t$CASES, decreasing = TRUE), ]

## Create subset with top 20
nc.zip.20 <- as_tibble(nc.zip.sp.t[1:20,])

## Create new field with for label
nc.zip.20$ZIPLABEL <- paste0(nc.zip.20$PO_NAME, 
                             " (", 
                             nc.zip.20$ZIP_CODE, 
                             ")")

## Create ggplot
ggplot(nc.zip.20,
       aes(x = reorder(ZIPLABEL, CASES),
           y = CASES)) + 
  geom_bar(stat = "identity") + 
  xlab("Zip Code") +
  ylab("COVID-19 Cases") +
  coord_flip() +
  theme_minimal()
```

******

### Choropleth Map

Now, we will use `tmap` to create a interactive choropleth map. We won't show the code, but we will show the map!


```{r echo = FALSE, warning = FALSE, message = FALSE, fig.width = 9, fig.height = 5}
## Put tmap in interactive mode
tmap_mode("view")

## Plot map
tm_shape(nc.zip.sp) + 
  tm_polygons("CASERATE", 
              style = "jenks", 
              palette = "YlOrRd",
              lwd = 0.25,
              border.col = "black",
              border.alpha = 0.5,
              title = "COVID-19 Case Rate<br>(per 1,000 people)") +
  tm_basemap("CartoDB.PositronNoLabels") +
  tm_view(set.view = 7)
```